---
title: "Exercises"
execute: 
  eval: true
  echo: true
---

```{r}
#| echo: false
params <- list()
params$choose_year <- "`r params$choose_year`"
```

## Data

Load the `gapminder` data:

```{r}
#| warning: false
library(gapminder)
```

## Exercises

### Exercise 1: Parameterized plots with `ggplot2`

1. Create a variable to define a year in the `gapminder` data that you want to plot.

<details>
<summary>Solution</summary>

```{r}
choose_year <- 1952
```

</details>

2. Edit the code below to use this new year variable.

```{r}
#| eval: false
library(dplyr)
library(ggplot2)
# Median life expectancy per continent
gapminder_1952 <- gapminder |>
  filter(year == 1952) |>
  group_by(continent) |>
  summarise(
    median_lifeExp = median(lifeExp)
  )
# Bar chart
ggplot(data = gapminder_1952) +
  geom_col(
    mapping = aes(
      x = median_lifeExp,
      y = reorder(continent, median_lifeExp)
    )
  ) +
  labs(
    x = "Median Life Expectancy (years)", y = NULL,
    title = "Life Expectancy in 1952"
  )
```

<details>
<summary>Solution</summary>

```{r}
#| message: false
library(dplyr)
library(ggplot2)
library(glue)

# Median life expectancy per continent
gapminder_year <- gapminder |>
  filter(year == choose_year) |>
  group_by(continent) |>
  summarise(
    median_lifeExp = median(lifeExp)
  )

# Bar chart
ggplot(data = gapminder_year) +
  geom_col(
    mapping = aes(
      x = median_lifeExp,
      y = reorder(continent, median_lifeExp)
    )
  ) +
  labs(
    x = "Median Life Expectancy (years)", y = NULL,
    title = glue("Life Expectancy in {choose_year}")
  )
```

</details>

3. Turn your code into a function that takes the data and year variable as arguments.

<details>
<summary>Solution</summary>

```{r}
barchart_year <- function(choose_year, data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      )
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
```

Check it works:

```{r}
barchart_year(2002)
```

</details>

4. **Bonus**: Add another argument to your function that allows a user to define the bar colour.

<details>
<summary>Solution</summary>

```{r}
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
```

Check it works:

```{r}
barchart_year(2002, fill = "purple")
```

</details>

5. **Bonus**: How could you improve your function?

<details>
<summary>Solution</summary>

* Not all year values are included in the data. Add a check that the value is a number that is included in the `gapminder` data. if it isn't, add a more useful message.
* Write some documentation!

</details>

### Exercise 2: Parameterized reports with Quarto

1. Edit the `report.qmd` file to add a parameter for year.

::: {.callout-warning title="report.qmd"}

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```

````

:::

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```
````

</details>

2. Add your plotting function to the quarto document, and pass the parameter into the plotting function. Check it renders.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````
</details>

3. Change the value of the parameter, and re-render.


### Exercise 3: Parameterized text with Quarto

1. Edit the sentence `...a plot of the median life expectancy in the year 1952 for each continent...` to use inline code to add the document parameter for the year. 

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

2. Use inline code for the title as well.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

3. Edit the figure caption to use inline code to add the document parameter for the year. 

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| message: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: !expr 'paste("Life expectancy in ", params$choose_year)'
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

### Exercise 4: Rendering multiple reports

1. Render a version of your report for the year `1972` using the `quarto_render()` function.

<details>
<summary>Solution</summary>

```{r}
#| eval: false
quarto_render(
  input = "report.qmd",
  execute_params = list(choose_year = 1972)
)
```

</details>

2. Render a version of your report for all years in the data.

<details>
<summary>Solution</summary>

```{r}
#| eval: false
walk(
  .x = unique(gapminder$year),
  .f = ~quarto_render(
    input = "report.qmd",
    output_file = glue("report_{.x}.html"),
    execute_params = list(
      choose_year = .x
    )
  )
)
```

</details>


### Exercise 5

Coming soon!

### Exercise 6

Coming soon!





